---
title: "map_colors_function"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Let's look at the poor old function

This is the function code from Barry's student

```{r}
# Old function
map.colors <- function (value,high.low,palette) {
  proportion <- ((value-high.low[1])/(high.low[2]-high.low[1]))
  index <- round ((length(palette)-1)*proportion)+1
  return (palette[index])
}
```


Now let's try to understand this function

```{r}
map.colors2 <- function(x, high.low = range(x), palette = cm.colors(100)) {
  
  ## Description: Map the values of the input vector 'x' to the input colors vector 'palett'
  
  # Determine where in the 'high.low' range our values of 'x' lie
  percent <- ((x - high.low[1])/(high.low[2] - high.low[1]))
  
  # Where in the 'palette' vector of colors is the 'percent'
  # note catch for 0 percent values to 1
  index <- round((length(palette) - 1) * percent) + 1
  
  return(palette[index])
}
```


## Working with our own functions

My first function

```{r}
add <- function(x, y=1) {
  # SUm the input x and y
  x + y
}
```

Let's test our first function

```{r}
add(c(1:5))
```

Does this work? No.
```{r}
#add(1, 3, 10)
```

Write my second function
```{r}
rescale <- function(x) {
  rng <- range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

Test my second function
```{r}
rescale(1:10)
```

```{r}
rescale2 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```


```{r}
#rescale2(c(1, 10, "string"))
```


```{r}
rescale3 <- function(x, na.rm = TRUE, plot = FALSE) {
  if(na.rm) {
    rng <- range(x, na.rm = na.rm)
  } else {
    rng <- rang(x)
  }
  print("Hello")
  
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  
  print("is it me you are looking for?")
  
  if (plot) {
    plot(answer, typ = "b", lwd = 4)
  }
  print("I can see it in...")
  return(answer)
}
```

```{r}
rescale3(c(1, 2, NA, 3, 10), plot = TRUE)
```



## Section 2B from Class 6

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain = "A", elety = "CA")
s2.chainA <- trim.pdb(s2, chain = "A", elety = "CA")
s3.chainA <- trim.pdb(s1, chain = "A", elety = "CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse = s1.chainA, typ = "l", ylab = "Bfactor")
plotb3(s2.b, sse = s2.chainA, typ = "l", ylab = "Bfactor")
plotb3(s3.b, sse = s3.chainA, typ = "l", ylab = "Bfactor")
```

```{r}
s1 <- read.pdb("4AKE")
s1
```

```{r}
s1.chainA <- trim.pdb(s1, chain = "A", elety = "CA")
s1.chainA
```

```{r}
s1.b <- s1.chainA$atom$b
s1.b
```

```{r}
plotb3(s1.b, sse = s1.chainA, typ = "l", ylab = "Bfactor")
```

# Q1 Returns brief summary from online of the protein (model, atoms, sequence)
# Q2 trim.pdb 'produces a new smaller PDB object, containing a subset of atoms, from a given larger PDB object', extracts small bit of the structure
# Q3 The secondary structure (beta strands and alpha helices) is shown in the margins. Remove the sse argument (default is NULL)
#Q4 One plot across the three (compare the three)



