---
title: "Class 8"
author: "Corrina Elder"
date: "April 27, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering example 

Let's make up some data for testing the `kmeans()` function.

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

Use the kmeans() function setting k to 2 and nstart=20
Inspect/print the results
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
    - cluster size?
    - cluster assignment/membership?
    - cluster center?
    
```{r}
#km$cluster
```
    
    
```{r}
#km$centers
```
    
Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points
      
```{r}
#plot(x, col = km$cluster, pch = 16)
#points(km$centers, col = "blue", pch = 15)
```
      
Q. Repeat for k=3, which one has the better total SS?

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

#Hierarchical clustering



First we need to calculate point (dis)similarity as the Euclidean distance between observations

```{r}
dist_matrix <- dist(x)
```

```{r}
class(dist_matrix)
```

Convert to matrix to see the structure of the distance matrix and find the dimensions (to view...View(as.matrix(dist_matrix)) )

```{r}
dim(as.matrix(dist_matrix))
```

The hclust() function returns a hierarchical clustering model

```{r}
hc <- hclust(d = dist_matrix)
```

```{r}
hc
```

```{r}
class(hc)
```

```{r}
plot(hc)
abline(h = 6, col = "tomato")
```

Let's 'cut' our tree to define our clusters

```{r}
grps <- cutree(hc, h = 6)
table(grps)
```

```{r}
plot(x, col = grps)
```

Try different cutting

```{r}
plot(x, col = cutree(hc, k = 4))
```


Try it out

# Step 1. Generate some example data for clustering

```{r}
x <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```


Step 2. Plot the data without clustering

```{r}
plot(x)
```


Step 3. Generate colors for known clusters
        (just so we can compare to hclust results)

```{r}
col <- as.factor( rep(c("c1","c2","c3"), each = 50) )
plot(x, col = col)
```


Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters

```{r}
d <- dist(x)
hc <- hclust(d)
plot(hc)
```

Let's cut into 2 groups

```{r}
grps <- cutree(hc, k = 3)
table(grps)
```

Plot the data colored by cluster
```{r}
plot(x, col = grps)
```

```{r}
plot(x, col = cutree(hc, k = 2))
plot(x, col = cutree(hc, k = 3))
```


# Q. How does this compare to your known 'col' groups?



# Principal Component Analysis

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names = 1)


head(mydata)
```

```{r}
#View(t(mydata))
```

```{r}
pca <- prcomp(t(mydata), scale = TRUE)

attributes(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2]) 
```

## Percent variance captured per PC
```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```

```{r}
barplot(pca.var.per, main = "Scree Plot",
 xlab = "Principal Component", ylab = "Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
colvec <- as.factor( substr( colnames(mydata), 1, 2) )
plot(pca$x[,1], pca$x[,2], col = colvec, pch = 16,
 xlab = paste0("PC1 (", pca.var.per[1], "%)"),
 ylab = paste0("PC2 (", pca.var.per[2], "%)"))
```

```{r}
mycols <- as.factor(substr(colnames(mydata), 1, 2))
plot(pca$x[,1], pca$x[,2], col = mycols, pch = 16,
     xlab = paste0("PC1 (",pca.var.per[1]," %)"), 
     ylab = paste0("PC2 (",pca.var.per[2]," %)"))

text(pca$x[,1] + 0.75, pca$x[,2], colnames(mydata))

```


#PCA of UK food data

```{r}
#par(mar=c(20, 4, 4, 2))
x <- read.csv("UK_foods.csv")
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
dim(x)

heatmap(as.matrix(x))
```

```{r}
pca <- prcomp( t(x) )
summary(pca)
```

```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
cumsum(v)
```


