---
title: "Bioinformatics Class 7"
author: "Corrina Elder"
date: "April 25, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Functions again

We can source any file of R code with the `source`() function

```{r}
source("http://tinyurl.com/rescale-R")
```

Let's make sure things are here

```{r}
ls()
```

Check our `rescale()` is working

```{r}
rescale(1:10)
```

```{r, eval = FALSE}
rescale( c(1:10, "string"))
```

Let's check if `rescale2()` does any better

```{r, eval = FALSE}
rescale2( c(1:10, "string"))
```


## Function for finding missing values in two datasets

Write a `both_na()` function to do this...

```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c(NA,3,NA,3, 4)

sum(is.na(x))
```

```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c(NA,3,NA,3, 4)

which(is.na(x))
```

```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c(NA,3,NA,3, 4)

which(!is.na(x))
```

```{r}
sum(is.na(x) & is.na(y))
```

My first function can start from this snippet

```{r}
both_na <- function(x, y) {
  sum( is.na(x) & is.na(y))
}
```

Test it!

```{r}
both_na(x, y)
```

Testing

```{r}
 x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x, y2)
```


```{r}
# x <-  c(NA, NA, NA)
#y1 <- c( 1, NA, NA)
#y2 <- c( 1, NA, NA, NA)

#both_na2(x, y2)
```

```{r}
x <- c( 1, 2, NA, 3, NA) 
y <- c(NA,3,NA,3, 4)

ans <- both_na3(x, y)
```

```{r}
ans$which
```

## And a last function that is useful

```{r}
x <- df1$IDs
y <- df2$IDs

x
y

intersect(x, y)
x %in% y        #for each element in x, check if it's in y
x[x %in% y] 
y[y %in% x]

cbind( x[ x %in% y ], y[ y %in% x ] ) #column bind into matrix
```

Make this into a function

```{r}
gene_intersect <- function(x, y) {
   cbind( x[ x %in% y ], y[ y %in% x ] )
}

x <- df1$IDs
y <- df2$IDs
gene_intersect(x, y)
```

Let's try with data.frome input rather than vectors

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

```{r}
gene_intersect4(df1, df2)
```

```{r}
merge(df1, df2, by="IDs")
```

















