---
title: "Lecture 11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Import our PDB stats CSV file and calculate percent structures by experimental method
```{r}
p <- read.csv("Data Export Summary.csv", row.names = 1)
```

Q1:
```{r}
percent <- p$Total / sum(p$Total) * 100
names(percent) <- row.names(p)
percent
```

What proportion of structures are protein?
```{r}
percent_protein <- sum(p$Proteins) / sum(p$Total) * 100
percent_protein
```


Q2: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?


## Section 2: Visualizing the HIV-1 protease structure

Use VMD

## Section 3: Using Bio3D

Load the bio3d package

```{r}
library(bio3d)
```

Read in our HIV-Pr structure

```{r}
pdb <- read.pdb("1hsg.pdb")
pdb
```

Q6: How many amino acid residues are there in this pdb object and what are the two non-protein residues?

198 amino acid residues
2 non-protein: HOH and MK1

Q7: What type of R object is pdb$atom? 

Dataframe

```{r}
str(pdb$atom)
```

```{r}
# Print a summary of the coordinate data in $xyz  
pdb$xyz
```

```{r}
dim(pdb$xyz)
pdb$xyz[1, atom2xyz(1:2)]
```

```{r}
# Select all C-alpha atoms (return their indices)  
ca.inds <- atom.select(pdb, "calpha") 
ca.inds
```

```{r}
# Print details of the first few selected atoms  
head( pdb$atom[ca.inds$atom, ] )

# And selected xyz coordinates  
head( pdb$xyz[, ca.inds$xyz] )
```

```{r}
# Select chain A 
a.inds <- atom.select(pdb, chain = "A") 
 
# Select C-alphas of chain A 
ca.inds <- atom.select(pdb, "calpha", chain = "A") 
 
# We can combine multiple selection criteria to return their intersection 
cab.inds <- atom.select(pdb, elety = c("CA","CB"), chain = "A", resno = 10:20)
```

Q8: Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing in VMD. Also write out a second separate PDB file for the ligand with residue name MK1

```{r}
inds.ligand <- atom.select(pdb, "ligand")
inds.protein <- atom.select(pdb, "protein")
inds.ligand
```

Check we have what we want
```{r}
pdb$atom[inds.ligand$atom,]
```

```{r}
head(pdb$atom[inds.protein$atom,])
```

```{r}
pdb.protein <- trim.pdb(pdb, inds = inds.protein)
pdb.ligand <- trim.pdb(pdb, inds = inds.ligand)
pdb.ligand
```

```{r}
write.pdb(pdb.ligand, file = "1hsg_ligand.pdb")

write.pdb(pdb.protein, file = "1hsg_protein.pdb")
```


## Section 4: Working with multiple PDB files

```{r}
# Download some example PDB files 
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")  
files <- get.pdb(ids, split = TRUE)
```

```{r}

# Extract and align the chains we are interested in  
pdbs <- pdbaln(files, fit = TRUE)
 
# Print to screen a summary of the 'pdbs' object  
pdbs
```


## Section 5: Example Application on Adenylate Kinase

```{r}
aa <- get.seq("1ake_A")
```

```{r}
# Blast or hmmer search
b <- blast.pdb(aa)
```

```{r}
# Plot a summary of search results
hits <- plot(b)
```

```{r}
head(hits$pdb.id)
```

```{r}
# Fetch PDBs
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
# Align structures
pdbs <- pdbaln(files)
# Vector containing PDB codes
ids <- basename.pdb(pdbs$id)
```

```{r}
# Draw schematic alignment
plot(pdbs, labels = ids)
```

```{r}
# Calculate sequence conservation
cons <- conserv(pdbs, method="entropy22") # SSE annotations
sse <- pdbs2sse(pdbs, ind=1, rm.gaps=FALSE)
# Plot conservation per residue
plotb3(cons, sse=sse, ylab="Sequence entropy")
```

```{r}
anno <- pdb.annotate(ids) 
print(unique(anno$source))
```


```{r}
# find invariant core
core <- core.find(pdbs)  
```

```{r}
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core) 
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)

pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```


```{r}
# Calculate RMSD
rd <- rmsd(pdbs)  
# Structure-based clustering
hc.rd <- hclust(dist(rd)) 
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)

```















